High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-07-02 inst/manuscript/si.Rmd
Manuscript and SI revised through Results (map accuracy section).

@@

## Assessing quality of composites
The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites. The imagery was evaluated by examining their Raster Foundry [@azaveaRasterFoundry2020] overlays within a `labeller` instance set up for the purpose.

@@

## Active learning process
### Labelling
The locations of training, training reference, and validation points are shown in Figure \ref{fig:trainval}. In AOI 3, the initial active learning cycle resulted in low accuracy because the northern part of the AOI shows low contrast between fields and the surrounding vegetation in the dry season. Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
```{r trainval, echo = FALSE, out.width="70%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected to initialize a second run for retrain AOI 3, and the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration. Grey background lines and show the boundaries and ID number of Ghana's agroecozones: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/si_training_validation_pts.png')
```

The distribution of data collection effort and scores from training reference sites by labelling team are shown in Figure \ref{fig:trainqual}. These results include those from re-labelling the initial training in Cluster 2 (AOIs 7-9, 12, 15). This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images. We replaced the image overlays in the instance for central-southeast (Figure \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019. The reprocessed labels were used to initially train the model for AOIs 9, 12, and 15.

### Validation
The differences in accuracy, AUC, and F1 between the active learning process and the random retraining at each iteration for AOIs 1, 8, and 15 (Figure \ref{fig:randomvactive}). Small differences due to random variations in the RandomForest models are the reason for the non-zero differences at iteration 0, when both models were trained with the same level set.   
```{r randomvactive, echo = FALSE, out.width="90%", fig.cap="The percent difference in performance metrics per iteration for AOIs 1, 8, and 15 (grey lines with numbers indicating AOI; the black line indicates average difference across the three AOIs) when comparing models trained using active learning versus those trained using randomly selected sites. Postive percentages indicate superior performance by active learning, negative percentages the inverse.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/si_random_vs_active.png')
```

## Map accuracy

### Reference label uncertainty

@@

### Categorical accuracy
In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{fig:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones (Figures \ref{fig:aoizones}B and Table \ref{tab:aezaccuracy}).

@@

>>>>

fmt <- function(x) format(x, big.mark = ",")
```

@@

## Image compositing 
### Variation in lengths of temporal window used
The typical window for compositing dry season imagery was December, 2018 to February, 2019, but in the cloudiest regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November, to ensure a sufficient density of images for compositing.

### Assessing composite quality
The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites. The imagery was evaluated by examining their Raster Foundry [@azaveaRasterFoundry2020] overlays within a `labeller` instance set up for the purpose.

@@

#### Map reference label uncertainty

@@

#### Accuracy assessment zones
The zones used to assess regional variation in map accuracy are illustrated in Figure \ref{fig:aoizones}.

@@

## Mapping cropland probabilities with active learning
```{r, echo=FALSE, message=FALSE, warning=FALSE}
initial <- train_val_sites$initial %>% as_tibble %>% group_by(Cluster) %>%
  count()
active <- train_val_sites$active %>% as_tibble %>% filter(usage == "train") %>% 
  group_by(aoi) %>% count()
typicaln <- raster::modal(initial$n) + raster::modal(active$n)
aoi10n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller10") %>% pull(n))
aoi14n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller14") %>% pull(n))
aoi15n <- (initial %>% filter(Cluster %in% 2) %>% pull(n)) + 
  (active %>% filter(aoi == "labeller15") %>% pull(n))
aoi3init <- initial %>% filter(!Cluster %in% 1:3) %>% pull(n)
aoi3n <- aoi3init + (active %>% filter(aoi == "labeller3") %>% pull(n))
```

The number of active learning iterations per AOI was three, for all but four AOIs. AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (\>83%) and showed little subsequent improvement. The models for these two AOIs were thus trained with `r aoi10n` - `r aoi14n` labels. AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI). In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S4A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.

### Labelling
The locations of training, training reference, and validation points are shown in Figure \ref{fig:trainval}. In AOI 3, the initial active learning cycle resulted in low accuracy because the northern part of the AOI shows low contrast between fields and the surrounding vegetation in the dry season. Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
```{r trainval, echo = FALSE, out.width="70%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected to initialize a second run for retrain AOI 3, and the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration. Grey background lines and show the boundaries and ID number of Ghana's agroecozones: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/si_training_validation_pts.png')
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("label_summary")  
toofew <- label_summary$nlabels %>% filter(n < 10) %>% pull(ID)
tsites <- label_summary$nlabels %>% filter(!ID %in% toofew) %>% 
  group_by(kml_type) %>% 
  summarize(tot = sum(n), mu = round(mean(n)), med = round(median(n)), 
            stdev = round(sd(n)), lower = min(n), upper = max(n))

# training rate
train_rate <- label_summary$nlabels %>% 
  tidyr::pivot_wider(names_from = kml_type, values_from = n) %>% 
  rename_all(funs(c("ID", "tr", "tref"))) %>% 
  mutate(trate = tref / tr) %>% filter(tr > 10)

tsitesc <- tsites %>% mutate(across(where(is.numeric), fmt))
# nvecs <- tsites$tot[1])
# muvecs <- fmt(tsites$mu[1])

The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core group of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each (Figure \ref{fig:trainqual}A). As each training and reference tasks was mapped by 4 separate labellers, `r tsitesc$tot[1]` sets of vector labels were made. Each labeller digitized an average of `r tsitesc$mu[1]`. Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites, with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped. The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)` (Figure \ref{fig:trainqual}B).

These results include those from re-labelling the initial training in Cluster 2 (AOIs 7-9, 12, 15). This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images. We replaced the image overlays in the instance for central-southeast (Figure \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019. The reprocessed labels were used to initially train the model for AOIs 9, 12, and 15.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("label_risk")
murisk <- label_risk$stats %>% group_by(usage) %>% 
  summarize(Risk = round(mean(Risk), 3)) %>% 
  tidyr::pivot_wider(names_from = usage, values_from = Risk)
mintrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
  filter(Risk == min(Risk))
maxtrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
  filter(Risk == max(Risk))
minvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
  filter(Risk == min(Risk))
maxvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
  filter(Risk == max(Risk))

# label clusters
cl1 <- label_risk$stats %>% filter(aoi %in% 1:6) %>% group_by(usage) %>% 
  summarize(Risk = round(mean(Risk), 3))
cl2 <- label_risk$stats %>% filter(aoi %in% c(7:9, 12, 15)) %>% 
  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
cl3 <- label_risk$stats %>% filter(aoi %in% c(10:11, 13:14, 16)) %>% 
  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))

# training clusters
cl2arisk <- label_risk$stats %>% filter(aoi %in% c("Cl2a")) %>% pull(Risk) %>% 
  round(3)
cl2brisk <- label_risk$stats %>% filter(aoi %in% c("Cl2b")) %>% pull(Risk) %>% 
  round(3)
```

The average Bayesian risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels. Risk was highest in the northen AOIs (AOIs 1-6; Figures S6-7), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S6-7), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`). Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north. Labeller experience also appeared to reduce risk, which we observed during a relabelling of the 500 initial random site in this cluster; the mean risk of the updated labels was `r cl2brisk`, compared to `r cl2arisk` for original labels.

### Model performance
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# random versus active data
data("iteration_metrics")

# process data: remove labeller from AOI and select AOIs 1-16
V1 <- c("Accuracy", "AUC", "F1")
aois <- c(paste0("labeller", c(1, 8, 15, "1r", "8r", "15r")))
imetrics <- iteration_metrics %>% 
  filter(Metric %in% V1 & AOI %in% aois) %>% 
  mutate(AOI = gsub("labeller", "", AOI)) %>% 
  mutate(type = case_when(
    AOI %in% c(1, 8, 15) ~ "active", 
    AOI %in% c("1r", "8r", "15r") ~ "random"
  )) %>% 
  mutate(AOI = gsub("r", "", AOI)) %>% 
  tidyr::pivot_wider(names_from = "type", values_from = "Score") %>% 
  mutate(delta = (active - random) / random * 100) 
imetrics_all <- imetrics %>% 
  filter(Iteration < 4) %>% 
  group_by(Iteration, Metric) %>% 
  summarize(active = mean(active), random = mean(random), 
            delta = mean(delta)) %>%
  ungroup() %>% mutate(AOI = "All") %>% 
  select(AOI, Iteration, Metric, active, random, delta)

ixl <- lapply(1:3, function(x) {
  ix <- imetrics_all %>% filter(Iteration == x) %>% pull(delta)
  ix <- round(ix, 1)
  names(ix) <- V1
  ix
})
names(ixl) <- as.roman(1:3)

i1acc <- imetrics %>% 
  filter(Iteration == 1 & Metric == "Accuracy") %>% 
  pull(delta) %>% round(2)
names(i1acc) <- imetrics %>% distinct(AOI) %>% pull()

u <- function(x) unname(x)
```

The differences in accuracy, AUC, and F1 between the active learning process and the random retraining at each iteration for AOIs 1, 8, and 15 (Figure \ref{fig:randomvactive}). Small differences due to random variations in the RandomForest models are the reason for the non-zero differences at iteration 0, when both models were trained with the same level set.

The lower accuracy of actively versus randomly trained models in earlier iterations was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:randomvactive}). In comparison, iteration 1 active learning accuracies were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively. Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.

```{r randomvactive, echo = FALSE, out.width="90%", fig.cap="The percent difference in performance metrics per iteration for AOIs 1, 8, and 15 (grey lines with numbers indicating AOI; the black line indicates average difference across the three AOIs) when comparing models trained using active learning versus those trained using randomly selected sites. Postive percentages indicate superior performance by active learning, negative percentages the inverse.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/si_random_vs_active.png')
```

## Map accuracy 
### Categorical accuracy
In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{fig:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones (Figures \ref{fig:aoizones}B and Table \ref{tab:aezaccuracy}).
